A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-05-14 manuscript.Rmd
added new sentence addressing the importance of reproducibility and openness for efficiency of knowledge accumulation

In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of data analyses in R [@baseR]. Reproducibility is the ability to draw identical conclusion from the same statistical analysis and the same data. In practice, reproducibility means to us that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones intially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992]. The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019]. Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase trustworthiness of empirical results [@epskamp2019rep].

@@

An alternative to Docker on Linux is Singularity [@kurtzerSingularityScientificContainers2017]. Just replace any docker calls with `singularity docker`, because singularity fully supports docker images. A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is currently only available on Linux.

@@

Finally, we conclude with some notes on project organization, which we think makes migrating projects to a reproducible workflow easier. The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses. Next, one needs to make sure that all files relevant for the analysis can be moved to another computer. To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder. A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.

>>>>

In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of data analyses in R [@baseR]. Reproducibility is the ability to draw identical conclusion from the same statistical analysis and the same data. In practice, reproducibility means to us that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones intially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992]. The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019]. Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase trustworthiness of empirical results [@epskamp2019rep]. Transparency and accessibility are  central scientific values and open, reproducible code will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].

@@

An alternative to Docker on Linux is Singularity [@kurtzerSingularityScientificContainers2017]. To use it, just replace any docker calls with `singularity docker`, because singularity fully supports docker images. A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is currently only available on Linux.

@@

Finally, we conclude with some notes on project organization, which we think makes migrating projects to a reproducible workflow easier. The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses. Next, one needs to make sure that all files relevant for the analysis can be moved to another computer. To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder because absolute paths are specific to a given computer. A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access. The here package solves two common issues with relative paths. First, it takes care of the fact that path separator characters vary across operating systems (typically, slash or backslash). Second, anchor points of relative paths may differ depending on context. For example, knitr interprets paths relative to the dynamic document whereas R has a current working directory that may change over the course of an R ssession. The `here` package provides consistent paths relative to the project directory. The following are three examples of refering to local files ranging from absolute paths with system-specific path separators (bad) to relative paths using the `here` package:
